### 2011-07-15 doc/theses/svirec/kap3.tex
chapter 4 finished

The repair of an R-XML tree which not satisfies $\mathcal{FD}$ set of functional dependencies is a pair of functions $\delta'$ and $\varrho'$ such that $RXT'$ tree composed of the original tree and the repair ($RXT' = \langle T, \delta' \cdot \delta, \varrho' \cdot \varrho \rangle$) weakly satisfies FD ($RXT' \models_w \mathcal{FD}$).\\
With a repair $\langle \delta, \varrho \rangle$ of R-XML tree and a set of labelled nodes $N$ of this tree, we denote with $Update_{\delta}(N)$ the set of nodes modified by $\delta$. Analogously, we denote $True_{\varrho}(N) = \{n \in N| \varrho(n) = true\}$ and $False_{\varrho}(N) =\{n \in N | \varrho(n) = false\}$.

>>>>

The repair of an R-XML tree which not satisfies $\mathcal{FD}$ set of functional dependencies is a pair of functions $\delta'$ and $\varrho'$ such that $RXT'$ tree composed of the original tree and the repair ($RXT' = \langle T, \delta' \cdot \delta, \varrho' \cdot \varrho \rangle$) weakly satisfies FD ($RXT' \models_w \mathcal{FD}$). The composition of the R-XML tree and the repair (i.e. composition of their $\delta$ and $\varrho$ functions) is defined as follows:

\begin{define}[Composition of $\delta$ functions]\label{compDelta}
The composition of two functions $\delta_1$ and $\delta_2$ asssociating values to leaf nodes is
\begin{displaymath}
    \delta_1 \cdot \delta_2(n) =
        \begin{cases}
            \delta_1(n) & \text{if $\delta_1(n)$ is defined over $n$,}\\
            \delta_2(n) & \text{otherwise (i.e. $\delta_1(n)$ is not defined over $n$).}
        \end{cases}
\end{displaymath}\qed
\end{define}

\begin{define}[Composition of $\varrho$ functions]\label{compVarrho}
The composition of two reliability functions $\varrho_1$ and $\varrho_2$ asssociating boolean value to nodes is $$\varrho_1 \cdot \varrho_2(n) = min(\varrho_1(n), \varrho_2(n)),$$ where to the boolean value $true$ (resp. $false$) is associated value 1 (resp. 0).
\qed
\end{define}

With a repair $\langle \delta, \varrho \rangle$ of R-XML tree and a set of labelled nodes $N$ of this tree, we denote with $Update_{\delta}(N)$ the set of nodes modified by $\delta$. Analogously, we denote $True_{\varrho}(N) = \{n \in N| \varrho(n) = true\}$ and $False_{\varrho}(N) =\{n \in N | \varrho(n) = false\}$.

### 2011-07-15 doc/theses/svirec/kap4.tex
chapter 4 finished

The proposed algorithm is based on the algorithm described in Section \ref{RepConstAnswers} presented in \cite{RepAndConsistentAnswer}. This algorithm was chosen beacause of simple representation of the XML data using a concept which corresponds with concept used in relational databases. Another reason is using besides modification of node value as an update operation also marking particular node information as unreliable, which can reveal forgotten inconsistencies in the data.

@@

The algorithm is splitted into three steps, where second and third step are repeated until all violations are repaired. In the first step, XML document and FDs are loaded, and XML tree with corresponding tree tuples are created. Next step of the algorithm computes repair groups containting repair candidates for FD violations. In the third step chosen repair candidate is applied to an XML tree. In Algorithm \ref{propAlgo} we can see the simplified process of repairing XML FD violations, which covers second and third step of the whole algorithm.

@@

After creating RW-XML tree from input XML data, set of tree tuples (defined in Definition \ref{treeTuple}) are constucted. Since a set $paths(D)$, containing all possible paths defined in DTD $D$, can be infinite (DTD can define recursive structure of elements), the actual content of $paths(D)$ is modified to reflect the current structure of the RW-XML tree. Since definition of a tree tuple says that answer of the path $p$ contains at most one element, our modification of $paths(D)$ has no effect on costructing tree tuples if a DTD defines some optional path which is not defined for the RW-XML tree $RWXT$ (the set $p(RWXT)$ is empty).

@@

With created RW-XML tree $RWXT$ and corresponding tree tuples from an input data, we can now procceed to compute repair groups. Repair group is a set of repair candidates, which repairs the same FD or modifies the same part of $RWXT$. Each repair candidate is a pair of functions $\delta$ and $\varrho$ ($\langle \delta, \varrho \rangle$), where either modifies value of an RW-XML tree node ($\delta$ is defined) or marking a node as unreliable ($\varrho$ is definded). The $\delta$ function of repair candidate is defined same as in the XML tree (Definition \ref{xmlTree}) and defines new value of the RW-XML tree node. Similarly, $\varrho$ function defines the node which is marked as unreliable.

@@

The function \texttt{computeRepairGroup}, responsible for creating the repair group, is shown in Algorithm \ref{repairGroupAlgo}. The function gets RW-XML tree, a functional dependency $F$, tuples $t_1$, $t_2$ and a set of repair groups $RGS$ and computes the repair group containig repair candidates as follows:

\begin{enumerate}
	\item First repair candidates are created using function $computeRepairs$ from Algorithm \ref{repAlgo} (line 1).Since R-XML tree is a special case of RW-XML tree where all weights are equal to zero, we can pass RW-XML tree as a parameter to this function.
    \item Next is checked if repair candidates intersect another candidades from existing repair groups (line 2)
    \begin{itemize}
    	\item If the candidates intersect with some group, they are added to this group (line 3)
        \item Otherwise new repair group containing repair candidates is created (line 5)
    \end{itemize}
\end{enumerate}

@@

Since each repair candidate consists of a set of nodes which are modified (marked as unreliable or has modified value), we can compute cost of each candidate. It is important to say, that in out approach, we prefer repair candidates that modify value of the nodes before those marking nodes as unreliable. Therefore we added the koeficient $k$ to the calculation of the repair candidate cost, so that the repair candidate modifying node value will have lower cost. The definition of repair candidate cost follows:

\begin{define}[Repair Candidate Cost]
Given an RW-XML tree $RXT$ and a repair candidate $R$, we define the cost of $R$ as:
$$cost(R) = \sum_{u \in Modified_\delta(R)} \omega(u) + \sum_{v \in Modified_\varrho(R)} \omega(v) \times k,$$
where $k$ is koeficient determining a priority of the repair candidate modifying value of node before that  marking node as unreliable.\qed
\end{define}

@@

\begin{example}\label{koefK}
Consider XML tree $XT$ and repair candidates from Example \ref{repairExample}. A default weight of all nodes $v$ of $XT$ is $\omega(v) = 0.5$. If the koeficient $k = 1$, cost of repair candidates would be as follows: $cost(R_1) = 0.5$, $cost(R_2) = 0.5$, $cost(R_3) = 2$, $cost(R_4) = 2$ and $cost(R_5) = 6$.
Let us assume, that a repair candidate with the lowest repair cost will be applied to $XT$. If we order costs of repair candidates from the lowest to the highest, you can see that candidates which have marked unreliable nodes are at the end (and will not be applied to $XT$).

Howerver, if we set the koeficient $k = 0.1$, costs of candidates would be sorted as follows: $cost(R_3) = 0.2$, $cost(R_4) = 0.2$, $cost(R_1) = 0.5$, $cost(R_2)  = 0.5$ and $cost(R_5) = 0.6$. You can see that order of costs has significantly changed and the cadidate which will be applied is $R_3$ (is marking nodes as unreliable).\qed
\end{example}

@@

The last step of our algorithm can be divided into two steps, namely selection of the repair candidate and subsequent application of the candidate to RW-XML tree.

\subsubsection{Selection of Repair Candidate}

The previous part of the algorithm computes repair groups containing repair candidates, which can repair violations of provided FDs. From this candidates, we must choose one that will be applied on RW-XML tree. Since we introduced a weight model to the repair group, we can use the weight of repair groups and the cost of repair candidates to choose suitable candidate. In our approach, we introduce two distinct algorithms: first that not involve the user to process of selection and the second one that use the user interaction.

@@

If the user has decided not to selecting repair candidates by himself, the algorithm goes into guess mode (staring on line 6). In this mode the algorithm checks all repair groups if some of it containts the repair candidate, that is sufficiently similar with some previously selected repair candidate (line 7-10). This similarity is checked by function \texttt{canBeUsedUserSelection} shown in Algorithm \ref{canUserSel}. If neither of candidates is sufficiently similar, the algorithm choose repair candidate with the lowest cost from the repair group with the lowest weight (line 15-16).

\begin{algorithm}[H]
\floatname{algorithm}{Function}
\caption{$selectRepairByUser(RGS, SR, t)$}
\label{selectUser}
\begin{algorithmic}[1]
\REQUIRE{\ \\
$RGS$ set of repair groups\\
$SR$: set of previously selected repair candidates by the user\\
$t$: modified nodes count treshold of previously selected repair candidates\\}
\ENSURE $R$: repair candidate

@@

The function \texttt{canBeUsedUserSelection} gets actual repair candidate $RC$, set of all previously selected repair candidates $SR$, modified nodes count treshold $t$ from interval $(0,1]$ of previously selected repair candidate and determine if $RC$ is sufficiently similar as some repair candidate from $SR$. To be similar with some previously selected repair candidate $S$, $RC$ must repair the same $FD$ as $S$, it must use the same update operation as $S$ and $\lambda(RC) \geq \lceil\lambda(S) \times t\rceil$ (line 2). Furthermore, if $\lambda(RC) = \lceil\lambda(S) \times t\rceil$, set of paths $PS(RC)$ must be a subset of $PS(S)$ (line 3). Otherwise, if $\lambda(RC) > \lceil\lambda(S) \times t\rceil$, there must exists a subset of $PS(S)$ with $\lceil\lambda(S) \times t\rceil$ elements that is subset of $PS(RC)$. In other words, without taking into account the treshold $t$, if a repair candidate $R_1$ modifies some nodes, the sufficiently similar repair candidate $R_2$ is such that modifies at least the same nodes as $R_1$ (when we say the same nodes we mean same paths representing those nodes). The treshold $t$ can reduce count of modified nodes of some previously selected repair candidate $R_1$, which means that sufficiently similar repair candidate needs to modify fewer nodes that are similar with nodes modified by $R_1$. In Example \ref{tresholdExample} is shown the usage of the treshold $t$.

\begin{algorithm}
\floatname{algorithm}{Function}
\caption{$canBeUsedUserSelection(SR, t, RC)$}
\label{canUserSel}
\begin{algorithmic}[1]
\REQUIRE{\ \\
$SR$: set of previously selected repair candidates by the user\\
$t$: suitability treshold $(0,1]$ of previously selected repair candidates\\
$RC$: actual repair candidate}
\ENSURE \TRUE\ if actual repair candidate is similar to some previous one.
\FORALL{$S$ in $SR$}
    \IF{$S$ repairs the same FD as $RC$ \AND $S$ use the same update operation as $RC$ \AND $\lceil\lambda(SC) \times t\rceil \leq \lambda(RC)$}
        \IF{$\lceil\lambda(SC) \times t\rceil = \lambda(RC)$ \AND $PS(RC) \subseteq PS(SC)$}
            \RETURN \TRUE
        \ENDIF
        \IF{$\lceil\lambda(SC) \times t\rceil < \lambda(RC)$ \AND there exists a subset $s$ of $PS(SC)$ with $\lceil\lambda(SC) \times t\rceil$ elements, that $s \subseteq$ $PS(RC)$}
            \RETURN \TRUE
        \ENDIF
   \ENDIF
\ENDFOR
\RETURN \FALSE
\end{algorithmic}
\end{algorithm}

\begin{example}\label{tresholdExample}
Consider repair candidate $R_1$ selected by user, that marks as unreliable two nodes and $PS(R_1) = \{/bib/book/title, /bib/book/title/text()\}$. Next consider repair candidate $R_2$ with $PS(R_2) = \{/bib/book/title/text()\}$, that the modified node is marked as unreliable and both $R_1$ and $R_2$ repairs the same $FD$. If treshold $t=1$, $R_2$ is not considered as sufficiently similar to $R_1$, because $\lceil\lambda(R_1) \times t\rceil > \lambda(R_2)$. However, if treshold is $t=0.5$, $\lceil\lambda(R_1) \times t\rceil = \lambda(R_2)$ and $PS(R_2) \subseteq PS(R_1)$, then $R_2$ is sufficiently similar to $R_1$.\qed
\end{example}

\subsubsection{Application of Repair Candidate}

>>>>

The proposed algorithm is based on the algorithm described in Section \ref{RepConstAnswers} presented in \cite{RepAndConsistentAnswer}. This algorithm was chosen because of simple representation of the XML data using a concept which corresponds with concept used in relational databases. Another reason is using besides modification of node value as an update operation also marking particular node information as unreliable, which can reveal forgotten inconsistencies in the data.

@@

The algorithm is split into three steps, where second and third step are repeated until all violations are repaired. In the first step, XML document and FDs are loaded, and XML tree with corresponding tree tuples are created. Next step of the algorithm computes repair groups containing repair candidates for FD violations. In the third step chosen repair candidate is applied to an XML tree. In Algorithm \ref{propAlgo} we can see the simplified process of repairing XML FD violations, which covers second and third step of the whole algorithm.

@@

After creating RW-XML tree from input XML data, set of tree tuples (defined in Definition \ref{treeTuple}) are constructed. Since a set $paths(D)$, containing all possible paths defined in DTD $D$, can be infinite (DTD can define recursive structure of elements), the actual content of $paths(D)$ is modified to reflect the current structure of the RW-XML tree. Since definition of a tree tuple says that answer of the path $p$ contains at most one element, our modification of $paths(D)$ has no effect on constructing tree tuples if a DTD defines some optional path which is not defined for the RW-XML tree $RWXT$ (the set $p(RWXT)$ is empty).

@@

With created RW-XML tree $RWXT$ and corresponding tree tuples from an input data, we can now proceed to compute repair groups. Repair group is a set of repair candidates, which repairs the same FD or modifies the same part of $RWXT$. Each repair candidate is a pair of functions $\delta$ and $\varrho$ ($\langle \delta, \varrho \rangle$), where either modifies value of an RW-XML tree node ($\delta$ is defined) or marking a node as unreliable ($\varrho$ is defined). The $\delta$ function of repair candidate is defined same as in the XML tree (Definition \ref{xmlTree}) and defines new value of the RW-XML tree node. Similarly, $\varrho$ function defines the node which is marked as unreliable.

@@

The function \texttt{computeRepairGroup}, responsible for creating the repair group, is shown in Algorithm \ref{repairGroupAlgo}. The function gets RW-XML tree, a functional dependency $F$, tuples $t_1$, $t_2$ and a set of repair groups $RGS$ and computes the repair group containing repair candidates as follows:

\begin{enumerate}
	\item First repair candidates are created using function $computeRepairs$ from Algorithm \ref{repAlgo} (line 1).Since R-XML tree is a special case of RW-XML tree where all weights are equal to zero, we can pass RW-XML tree as a parameter to this function.
    \item Next is checked if repair candidates intersect another candidates from existing repair groups (line 2)
    \begin{itemize}
    	\item If the candidates intersect with some group, they are added to this group (line 3)
        \item Otherwise new repair group containing repair candidates is created (line 5)
    \end{itemize}
\end{enumerate}

@@

Since each repair candidate consists of a set of nodes which are modified (marked as unreliable or has modified value), we can compute cost of each candidate. It is important to say, that in out approach, we prefer repair candidates that modify value of the nodes before those marking nodes as unreliable. Therefore we added the coefficient $k$ to the calculation of the repair candidate cost, so that the repair candidate modifying node value will have lower cost. The definition of repair candidate cost follows:

\begin{define}[Repair Candidate Cost]
Given an RW-XML tree $RXT$ and a repair candidate $R$, we define the cost of $R$ as:
$$cost(R) = \sum_{u \in Modified_\delta(R)} \omega(u) + \sum_{v \in Modified_\varrho(R)} \omega(v) \times k,$$
where $k$ is coefficient determining a priority of the repair candidate modifying value of node before that  marking node as unreliable.\qed
\end{define}

@@

\begin{example}\label{koefK}
Consider XML tree $XT$ and repair candidates from Example \ref{repairExample}. A default weight of all nodes $v$ of $XT$ is $\omega(v) = 0.5$. If the coefficient $k = 1$, cost of repair candidates would be as follows: $cost(R_1) = 0.5$, $cost(R_2) = 0.5$, $cost(R_3) = 2$, $cost(R_4) = 2$ and $cost(R_5) = 6$.
Let us assume, that a repair candidate with the lowest repair cost will be applied to $XT$. If we order costs of repair candidates from the lowest to the highest, you can see that candidates which have marked unreliable nodes are at the end (and will not be applied to $XT$).

However, if we set the coefficient $k = 0.1$, costs of candidates would be sorted as follows: $cost(R_3) = 0.2$, $cost(R_4) = 0.2$, $cost(R_1) = 0.5$, $cost(R_2)  = 0.5$ and $cost(R_5) = 0.6$. You can see that order of costs has significantly changed and the candidate which will be applied is $R_3$ (is marking nodes as unreliable).\qed
\end{example}

@@

The last step of our algorithm can be divided into two parts, namely selection of the repair candidate and subsequent application of the candidate to RW-XML tree.

\subsubsection{Selection of the Repair Candidate}

The previous step of the algorithm computes repair groups containing repair candidates, which can repair violations of provided FDs. From this candidates, we must choose one that is applied on RW-XML tree. Since we introduced a weight model to the repair group, we can use the weight of repair groups and the cost of repair candidates to choose suitable candidate. In our approach, we introduce two distinct algorithms: first that not involve the user to process of selection and the second one that use the user interaction.

@@

If the user has decided not to selecting repair candidates by himself, the algorithm goes into guess mode (staring on line 6). In this mode the algorithm checks all repair groups if some of it contains the repair candidate, that is sufficiently similar with some previously selected repair candidate (line 7-10). This similarity is checked by function \texttt{canBeUsedUserSelection} shown in Algorithm \ref{canUserSel}. If neither of candidates is sufficiently similar, the algorithm choose repair candidate with the lowest cost from the repair group with the lowest weight (line 15-16).

\begin{algorithm}[H]
\floatname{algorithm}{Function}
\caption{$selectRepairByUser(RGS, SR, t)$}
\label{selectUser}
\begin{algorithmic}[1]
\REQUIRE{\ \\
$RGS$ set of repair groups\\
$SR$: set of previously selected repair candidates by the user\\
$t$: modified nodes count threshold of previously selected repair candidates\\}
\ENSURE $R$: repair candidate

@@

The function \texttt{canBeUsedUserSelection} gets actual repair candidate $RC$, set of all previously selected repair candidates $SR$, modified nodes count threshold $t$ from interval $(0,1]$ of previously selected repair candidate and determine if $RC$ is sufficiently similar as some repair candidate from $SR$. To be similar with some previously selected repair candidate $S$, $RC$ must repair the same $FD$ as $S$, it must use the same update operation as $S$ and $\lambda(RC) \geq \lceil\lambda(S) \times t\rceil$ (line 2). Furthermore, if $\lambda(RC) = \lceil\lambda(S) \times t\rceil$, set of paths $PS(RC)$ must be a subset of $PS(S)$ (line 3). Otherwise, if $\lambda(RC) > \lceil\lambda(S) \times t\rceil$, there must exists a subset of $PS(S)$ with $\lceil\lambda(S) \times t\rceil$ elements that is subset of $PS(RC)$. In other words, without taking into account the threshold $t$, if a repair candidate $R_1$ modifies some nodes, the sufficiently similar repair candidate $R_2$ is such that modifies at least the same nodes as $R_1$ (when we say the same nodes we mean same paths representing those nodes). The threshold $t$ can reduce count of modified nodes of some previously selected repair candidate $R_1$, which means that sufficiently similar repair candidate needs to modify fewer nodes that are similar with nodes modified by $R_1$. In Example \ref{tresholdExample} is shown the usage of the threshold $t$.

\begin{algorithm}
\floatname{algorithm}{Function}
\caption{$canBeUsedUserSelection(SR, t, RC)$}
\label{canUserSel}
\begin{algorithmic}[1]
\REQUIRE{\ \\
$SR$: set of previously selected repair candidates by the user\\
$t$: suitability threshold $(0,1]$ of previously selected repair candidates\\
$RC$: actual repair candidate}
\ENSURE \TRUE\ if actual repair candidate is similar to some previous one.
\FORALL{$S$ in $SR$}
    \IF{$S$ repairs the same FD as $RC$ \AND $S$ use the same update operation as $RC$ \AND $\lceil\lambda(SC) \times t\rceil \leq \lambda(RC)$}
        \IF{$\lceil\lambda(SC) \times t\rceil = \lambda(RC)$ \AND $PS(RC) \subseteq PS(SC)$}
            \RETURN \TRUE
        \ENDIF
        \IF{$\lceil\lambda(SC) \times t\rceil < \lambda(RC)$ \AND there exists a subset $s$ of $PS(SC)$ with $\lceil\lambda(SC) \times t\rceil$ elements, that $s \subseteq$ $PS(RC)$}
            \RETURN \TRUE
        \ENDIF
   \ENDIF
\ENDFOR
\RETURN \FALSE
\end{algorithmic}
\end{algorithm}

\begin{example}\label{tresholdExample}
Consider repair candidate $R_1$ selected by user, that marks as unreliable two nodes and $PS(R_1) = \{/bib/book/title, /bib/book/title/text()\}$. Next consider repair candidate $R_2$ with $PS(R_2) = \{/bib/book/title/text()\}$, that the modified node is marked as unreliable and both $R_1$ and $R_2$ repairs the same $FD$. If threshold $t=1$, $R_2$ is not considered as sufficiently similar to $R_1$, because $\lceil\lambda(R_1) \times t\rceil > \lambda(R_2)$. However, if threshold is $t=0.5$, $\lceil\lambda(R_1) \times t\rceil = \lambda(R_2)$ and $PS(R_2) \subseteq PS(R_1)$, then $R_2$ is sufficiently similar to $R_1$.\qed
\end{example}

\subsubsection{Application of the Repair Candidate}

In this part, the selected repair candidate is finally applied to the RW-XML tree. If after this part the RW-XML tree not violates any of FDs, the whole repair algorithm ends at this point, otherwise repair groups are regenerated and the selection of the repair candidate part take place.

To apply the selected repair candidate $R$ to the RW-XML tree $RXT$ means, that we compose $\delta$ and $\varrho$ functions of $R$ with the corresponding functions of R-XML tree contained in $RXT$. This compositions are defined in Definition \ref{compDelta} and \ref{compVarrho}.

After application of the repair candidate, some of RW-XML tree nodes could become unreliable, which can lead to the situation that some of tree tuples are not anymore considered as a tuple (is no longer a maximal subtree). Therefore, we need before regeneration of repair groups also check all tree tuples, if they satisfies definition of tree tuple.
